\chapter{Personal Reflection}\label{sec:personalreflection}

Personally, this bachelor thesis research has been instructive and beneficial in many different ways. Most importantly, I have had the opportunity to dedicate my entire time to learn about the subject of quantum information and, specifically, quantum machine learning in detail. Since these subjects are not taught within the curriculum of the Maastricht Science Programme (MSP) it was especially amazing to challenge myself to these notoriously difficult subjects within the intersection of quantum physics and computer science. In doing so, I have learned more about the methods of theoretical physics and gained additional experience in scientific programming with Octave, Python and F\#. Without prior knowledge of F\#, I was able to learn how to simulate quantum computations and quantum machine learning algorithms using the quantum simulation toolsuite Liqui$\ket{}$. Furthermore, I was able to use the first cloud-based quantum computer, the so-called IBM Quantum Experience, publicly released in May 2016 by IBM.

Alongside my research, I had the possibility to attend many great lectures, seminars and three conferences which were all generously funded by the Centre for Quantum Technology. This enabled me to get to know many renowned scientists working in the fields of quantum information, quantum machine learning, open quantum systems, quantum optics and quantum cryptography. Furthermore, I was provided with the opportunity to present my research at the 4\textsuperscript{th} South African conference for quantum information processing, communication and control in Cape Town in late November 2016. This constituted a big step with respect to my academic career and a great way of putting my acquired presentation skills to practice.

However, in retrospective I feel that I could have been more productive in office by structuring my work days more diligently. For example, the first one and a half months were mostly spent on reading research paper and text books to acquire the theoretical foundation for my research and little time was used to practice F\# within the Liqui$\ket{}$ framework. A possible solution for the future would be to divide each day into two parts: e.g. spending the morning with reading papers and books and the afternoon on practical work and programming. Besides this, I feel like I should have communicated my progress more with my internal supervisor Dr. Birembaut at the MSP.

In conclusion, the bachelor thesis research has showed me the importance of interdisciplinarity since the field of quantum information requires the understanding of concepts in computer science as well as quantum mechanics. Fortunately, MSP's liberal education does not only focus on textbook exercises and lectures but mostly teaches us how to systematically approach unknown topics and tackle problems therein. It was great to see that this enabled me to venture into a previously unfamiliar field and learn the necessary skills to conduct meaningful research. The newly acquired skills will certainly be useful for my planned masters in the field of quantum information.

%During the thesis work, I realized that encountering problem after problem is at the core of research but that one can resolve most of them when just showing enough dedication.

%The Centre for Quantum Technology at the University of KwaZulu-Natal in South Africa has been a wonderful host for my Bachelor thesis research. My supervisor Prof. Francesco Petruccione has become a mentor and friend and provided me with great opportunities for personal and academic growth. Furthermore, my collaborator Maria Schuld has shared vasts amount of knowledge, tricks and ideas with me and has always been a great help during my research work. Alongside my research, I had the possibility to attend many great lectures, seminars and three conferences which were all generously funded by the Centre for Quantum Technology. This enabled me to get to know many renowned scientists working in the fields of quantum information, quantum machine learning, open quantum systems, quantum optics and quantum cryptography. Attending the Quantum Machine Learning Workshop at the Dolphin Coast in July 2016 provided me with a broad overview of the field of quantum machine learning and was the first time that I ever attended a scientific conference. In November 2016, I was provided with the opportunity to present my research at the 4\textsuperscript{th} South African conference for Quantum Information Processing, Communication and Control in Cape Town. This opened the possibility for a collaboration with an experimental research group in Israel working on quantum computation with trapped ions. In January 2017, I am invited to the NiTheP Chris Engelbrecht Quantum Machine Learning Summer School where I will give three workshop sessions on quantum machine learning using the Liqui$\ket{}$ framework and the IBM Quantum Experience.